

Step 1: Move all data onto a single Excel tab. Add a column for church location and change column names to single-word variables. Don't forget to pay attention to bulk meal requests. Add a Unique_ID column at the front. Save results as a separate .csv document.

Step 2: Import address data. 
```{r}
library(tidyr)
library(dplyr)
library(XLConnect)

df <- tbl_df(read.csv("BranchesThanksgivingClientListMaster.csv", header = TRUE))

locations <- unique(df$Church_Name)
```

Step 3: Separate data into locations, assign a route to each address
```{r}
location_count <- 1
df_route_assign <- list()
for (location in locations){
  df_location <- df %>% 
    filter(Church_Name == location) %>% 
    arrange(Client_Zip) %>% 
    mutate(Route = 0) %>% 
    mutate(Order = row_number(Client_Zip))

  items <- unique(df_location$Order)
  
  current_route <- 1
  df_location$Route[1] <- current_route
  
  for (item in items[2:length(items)]) {
    
    # Calculate the number of meals and the stops on the current route
    route_summary <- df_location %>% 
                        filter(Route == current_route) %>% 
                        summarize(Meals = sum(Client_Meals), Stops = n())
    
    if((df_location$Client_Zip[item] == df_location$Client_Zip[item - 1]) & # Zip codes match
        (route_summary$Meals[1] + df_location$Client_Meals[item] <= 20) & # Meals can't exceed 20
        (route_summary$Stops[1] + 1 <= 4)){ # Stops can't exceed 4
            df_location$Route[item] <- current_route
    }
    else if ((route_summary$Meals[1] + df_location$Client_Meals[item] <= 20) &
             (route_summary$Stops[1] <= 2)){
            df_location$Route[item] <- current_route
    }
    else {
            # Set up the new route
            current_route <- current_route + 1
            df_location$Route[item] <- current_route
    } # End of else statement in if-else decision tree
    
  } # End of item for-loop
  
  # Save the data to inspect route assignments
  df_route_assign[[location_count]] <- df_location
  location_count <- location_count + 1
  
} # End of location for-loop
```
Make sure to inspect the routes before moving on to step 4

Step 4: Create a spreadsheet for each loop
```{r}
for (loc in 1:length(locations)) {

  df_loc <- df_route_assign[[loc]]
  routes <- unique(df_loc$Route)
  location <- locations[loc]
  
  for (route in routes) {
      
      final_route_summary <- df_loc %>% 
                        filter(Route == route) %>% 
                        summarize(Meals = sum(Client_Meals), Stops = n())
      wb <- loadWorkbook("Route_Template.xlsx")

      style.title  <- getCellStyle (wb , "my_header")
      style.normal <- getCellStyle (wb , "my.table")

      writeWorksheet(wb,route, sheet = 1 ,startRow = 2, startCol = 2, header = F)
      writeWorksheet(wb,location, sheet = 1 ,startRow = 3, startCol = 2, header = F)
      writeWorksheet(wb,final_route_summary$Meals[1], sheet = 1 ,startRow =2, startCol=9, header= F)
      writeWorksheet(wb,final_route_summary$Stops[1], sheet = 1 ,startRow =3, startCol=9, header= F)

      setCellStyle(wb, sheet=1, row= 2:3, col=1:9, cellstyle = style.title)

      route_final <- df_loc %>%
                      filter(Route == route) %>%
                      mutate(New_Order = row_number(Order))
      route_item <- unique(route_final$New_Order)

      for (item in route_item) {
        # Populate the worksheet
        writeWorksheet(wb,route_final$Client_Name[item],
                       sheet = 1 ,startRow = 5 + item, startCol = 1, header = F)
        writeWorksheet(wb,route_final$Client_Address[item],
                       sheet = 1 ,startRow = 5 + item, startCol = 2, header = F)
        writeWorksheet(wb,route_final$Client_Apt[item],
                       sheet = 1 ,startRow = 5 + item, startCol = 3, header = F)
        writeWorksheet(wb,route_final$Client_City[item],
                       sheet = 1 ,startRow = 5 + item, startCol = 4, header = F)
        writeWorksheet(wb,route_final$Client_Zip[item],
                       sheet = 1 ,startRow = 5 + item, startCol = 5, header = F)
        writeWorksheet(wb,route_final$Client_Directions[item],
                       sheet = 1 ,startRow = 5 + item, startCol = 6, header = F)
        writeWorksheet(wb,route_final$Client_Num[item],
                       sheet = 1 ,startRow = 5 + item, startCol = 7, header = F)
        writeWorksheet(wb,route_final$Client_Language[item],
                       sheet = 1 ,startRow = 5 + item, startCol = 8, header = F)
        writeWorksheet(wb,route_final$Client_Meals[item],
                       sheet = 1 ,startRow = 5 + item, startCol = 9, header = F)

        # Set the cell style
        setCellStyle(wb, sheet=1, row= 5 + item, col=1:9, cellstyle = style.normal)
        setCellStyle(wb, sheet=1, row= 2, col=2, cellstyle = style.title)
        setCellStyle(wb, sheet=1, row=3, col=9, cellstyle = style.title)
      }
      
      saveWorkbook(wb, paste(location, " - ", route, ".xlsx", sep=""))
      
  } # End of route for-loop
            
} # End of location for-loop
```

Step 5: Validate the Excel outputs

Step 6: Print out route summaries to create the routing workbook for each location
```{r}
for (loc in 1:length(locations)) {
  
  route_stats <- df_route_assign[[loc]] %>% 
                  group_by(Route) %>% 
                  summarise(Church_Name = first(Church_Name), Meals = sum(Client_Meals), Stops = n())
  
  
  print(route_stats)

}
```


